---
title       : What R you on about?
subtitle    : Introduction to the wonderful world of R
author      : Bulukani Mlalazi
job         : Research Scientist
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## First up: A note to tha Skrybz!!

A small favour to those taking notes..

The only link that you really need to right down is here:

<b>https://github.com/drbulu/intro_to_r</b>

The entire presentation has been uploaded:  
1. HTML:  <b style="color:green">index.html</b>  
2. PDF: <b style="color:purple">create e.g. by printing in browser</b>  

--- .class #id 

## Presentation Outline

What is R?  

Where is R used?  

Who R users?  

Why use R?  

How does R work?  

--- .class #id 

## What is R?

1. R is a dialect/derivative of the language S  
2. Uses of R:  

  a) Data analysis  
  b) Data mining  
  c) Data Visualisation  

3. Some characteristics of R: 

  a) Open Source: Well supported platform  
  b) Extensible: Via custom packages  

More information:  

http://en.wikipedia.org/wiki/R_%28programming_language%29

--- .class #id 

## Where is R used?

Some Application Examples:  

<table>
  <tr>
    <td>Graphics and Visualizations</td>
    <td><b>HighThroughput Genomics</b></td>
  </tr>
    <td>Environmental Studies</td>
    <td><b>Medical Imaging</b></td>
  <tr>
    <td>Clinical Trials</td>
    <td><b>Experimental Design</b></td>
  </tr>
  <tr>
    <td>Graphics and Visualizations</td>
    <td><b>High Performance Computing</b></td>
  </tr>
  <tr>
    <td>Multivariate Statistics</td>
    <td><b>Natural Language Processing</b></td>
  </tr>  
</table>  
<br/>
More examples:  
  http://www.revolutionanalytics.com/applications-r  

--- .class #id 

## Who R users?

R is designed to be used by people can be used by is suitable for skills lie across the entire programming skill spectrum from: 
<ul>
  <li>non-programmers (<b>n00bs</b>)</li>  
  <li><b style="color:orange" >everyone in between</b></li>
  <li>the highly skilled (<b>1337</b>)</li>  
</ul>
<b>Examples of target users include:</b>  
1) scientists (of all kinds):  
<ul>
  <li>a) life scientists</li>
  <li>b) data scientists</li>
  <li>c) social scientists</li>
</ul>
2) hobbyists and script junkies  
<div align="center" ><b style="color:red" >in short... YOU!</b></div>

--- .class #id 

## Who R users?

R even has a journal..

  <img class=center src=./assets/img/rjournal_pic.png height='400px' /><br/>

  http://journal.r-project.org/

--- .class #id 

## Why use R?

### [KDNuggets 2014][link] Data Analysis sofware poll (Top 10)

1. RapidMiner, 44.2% share (39.2% in 2013)
2. <b>R</b>, <b>38.5%</b> ( 37.4% in 2013)
3. Excel, 25.8% ( 28.0% in 2013)
4. SQL, 25.3% ( na in 2013)  
5. Python, 19.5% ( 13.3% in 2013)  
6. Weka, 17.0% ( 14.3% in 2013)
7. KNIME, 15.0% ( 5.9% in 2013)
8. Hadoop, 12.7% ( 9.3% in 2013)
9. SAS base, 10.9% ( 10.7% in 2013)
10. Microsoft SQL Server, 10.5% (7.0% in 2013) 

[link]: http://www.kdnuggets.com/polls/2014/analytics-data-mining-data-science-software-used.html

--- .class #id 

## Why use R?
 
[RedMonk][redmonk] Programming Language Rankings (June 2014)  
   
<img class=center src=./assets/img/redmonk_popular.png height='300px' /><br/>

Image Source: [revolutionanalytics.com][revanalytics1]  

<span style="color:green;" >Yet *another* nice comparison can be found </span> [here][r4stats]!

[revanalytics1]: http://blog.revolutionanalytics.com/2014/06/r-popularity-update.html
[redmonk]: http://redmonk.com/sogrady/2014/06/13/language-rankings-6-14/
[r4stats]: http://r4stats.com/articles/popularity/

--- .class #id 

## Why use R?

<div align="center" ><b style="color:green" >Example 1: ggplot2</b></div>

[ggplot2][plotting_maddly] is a powerful data visualisation package for R.

<img class=center src=./assets/img/pie_chart31.png height='250px' />

[Example Source][pieface]

[pieface]: http://www.r-bloggers.com/pie-charts-in-ggplot2/
[plotting_maddly]: http://ggplot2.org/  

--- .class #id 

<div align="center" ><b style="color:green" >Example 1: some ggplot2 pictures</b></div>

<table>
  <tr>
    <th>
      <a href="http://www.r-bloggers.com/u-s-unemployment-data-animated-choropleth-maps/" >Unemployment in Greenland! (USA)
      <a/>
    </th>
    <th>
      <a href="http://www.r-bloggers.com/bp-oil-and-gas-recovery/" >Oil and Gas chart<a/>
    </th>
  </tr>
  <tr>
    <td>
      <img class=center src=./assets/img/2010_unemployment.png width='400px'/>
    </td>
    <td>
      <img class=center src=./assets/img/BP_oil_and_gas_recovery.png width='400px'/>
    </td>
  </tr>
</table>

--- .class #id 

<div align="center" ><b style="color:green" >Example 1: more ggplot2 pictures</b></div>

<table style="border: 1px solid black;border-collapse: collapse;">
  <tr>
    <th>
      <a href="http://www.r-bloggers.com/displaying-data-using-level-plots/" >Using Level Plots
      <a/>
    </th>
    <th>
      <a href="http://www.r-bloggers.com/summarising-data-using-histograms/" >Using Histograms<a/>
    </th>
  <tr>
  <tr>
    <td>
      <img class=center src=./assets/img/levelplot-ggplot2-300x300.jpg />
    </td>
    <td>
      <img class=center src=./assets/img/histogram-ggplot2-300x300.jpg />
    </td>
  </tr>
</table>

--- .class #id 

## Why use R?

<div align="center" ><b style="color:green" >Example 2: Slidify</b></div>

<img class=center src=./assets/img/slidify_pic.png height='300px' /><br/>

link: *http://slidify.org/*

This presentation was made using slidify!

--- .class #id 

## Why use R?

<div align="center" ><b style="color:green" >Example 3: Bioinformatics</b></div>

<img class=center src=./assets/img/logo_bioconductor.gif height='200px' /><br/>

link: *http://www.bioconductor.org/*

--- .class #id 

## Why use R?

### My experience

1) R is a fully fledged environment with tons of features

2) I prefer to use R over Excel whenever I can… it is usually just a matter of working out how to get the desired outcome in R  

3) R is much better for data analysis automation IMHO  

4) However, don’t get carried away. Every system has pros and cons… choose what works best for your 

#   a) data analysis needs   
#   b) IT skill level   


--- .class #id 

## How does R work?

A brief example of "Vectorised" operations

```{r}
  x = c(1:5);
  x
  print(x)
```

For more info:  
http://cran.r-project.org/doc/manuals/r-release/R-intro.html#Vector-arithmetic

--- .class #id 

## How does R work?

a little bit more depth

```{r}
  y <- x + 1; y
  length(y)
  y[2]  
```

--- .class #id 

## How does R work?

just a little bit more depth

```{r}
  z <- (x * 10) + y ; z
  z[length(z)]
```

--- .class #id 

## How does R work?

### Caution: Baby steps alert!

### This is a very basic look at how R works...  

### If you want to know more:  

<h2 style="color:green">1) Explore</h2>

<h2 style="color:red">2) lobby for follow up workshopz :)</h2>

--- .class #id

## How to get started?

Data Science (John Hopkins University):  
1. Coursera Platform: https://www.coursera.org/specialization/jhudatascience/1/overview  
2. Github: https://github.com/DataScienceSpecialization/courses  

Quick-R: http://www.statmethods.net/index.html  

R bloggers: http://www.r-bloggers.com/  

Markdown:  
1. http://daringfireball.net/projects/markdown/syntax  
2. https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet  

KnitR basic tutorial:  http://kbroman.org/knitr_knutshell/

etc...

For more information see <b style="color:red;text-decoration:underline;" >Googlz of the interwebz!</b>

--- .class #id 

## What R you going to do next?

<b style="color:red">i.e. where to from here?</b>  

a) Follow the links  

b) Ask someone who might know:  
    i) the interwebz  
        * google  
        * Stack Overflow (1337 h3lp)  
    ii) friendly p33pz  

c) Play around... lots of training datasets available
 
d) Apply it to your work ... one simple task at a time 

--- .class #id


<div align="center" ><h2 style="color:red;text-decoration:underline;" >Thank you!</h2></div>